### 2020-12-08 paper.Rmd
Add files via upload

---
title: "DSCI 445 - Algal Blooms"
author: "Dawson Eliasen, Boston Lee, Miles Austin"
date: "12/7/2020"
output:
  pdf_document: default
---

  Algae blooms are becoming more relevant in today’s world, due to the increasing 
climate change. Scientific research has shown that algae blooms are happening at
an increasing frequency, with a higher geographic distribution than we had 
previously seen, which is especially notable in high elevation reservoirs. Not 
only are algae blooms occurring more frequently, harmful algal blooms - also 
known as HABS - are also occurring more frequently.

  Even though algae are natural and part of a healthy freshwater ecosystem, the
algae can grow to an extent that it can be harmful to the ecosystem – creating 
anoxic conditions for example, otherwise known as “dead zones” for the aquatic 
fauna. The HABs can produce toxins that can be harmful to not only the aquatic 
ecosystem, but to humans and other animals, such as dogs. Both algae blooms and 
HABs have a wide range of impacts that can not only be local, but they can also
have interstate impacts. This is especially important to note because Lake Erie,
in 2014, had a HAB that was incredibly impactful to the local economy and 
residents. The HAB left half-a-million people without clean water to drink and 
roughly $65 million in economic losses over a two day span of time.

  With the increasing use of reservoirs for municipalities, the quality of the 
water is very important. Although the water gets treated before it gets to the 
tap, the quality of the water as it reaches the water treatment plant is very 
important regarding the cost to treat the water for municipal use, and algal 
blooms impact the quality of the water dramatically. Most municipal water is 
monitored, but not all of the data is publicly available and some isn’t 
monitored in situ - otherwise known as collected by hand. However, Landsat is a
series of satellites that image the Earth surface, in repeated intervals of 14 
days; that is to say Fort Collins gets re-imaged every 14 days. Landsat 5 dates
back to 1984, and is no longer taking images, while landsat 7 (1999) and landsat
8 (2013) still record images today. The landsat imagery allows for a larger 
sample of data to look at the trends of algal blooms in Colorado, in comparison 
to the lack of public in situ data available. Figure 1 shows reservoirs - only 
those greater than 0.01 km² -  that have both in situ and satellite imagery in 
blue, while the reservoirs in red indicate the reservoirs with just the landsat 
imagery available. It is easy to see the lack of reservoir data we have 
available to us, which is why utilizing machine learning to try and predict 
algae blooms from just remotely sensed imagery would be an incredibly useful
method of determining water quality in reservoirs.

Methods:
How we got the data
+/- day merge
Tried 3 models
Tried 5 variables
Tried day threshold
Train/test split? (⅓ was held out)
Experimental variables were compared using 5-fold cross validation

Results: